The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-11-28 article_final/main.tex
fixed a ffiliations

\author{\Name{Alexey Kravatskiy}
       \Email{kravtskii.aiu@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)
       \AND
       \Name{Ivan Kozyrev}
       \Email{kozyrev.in@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)\\
       Marchuk Institute of Numerical Mathematics
       \AND
       \Name{Nikolai Kozlov}
       \Email{kozlov.na@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)
       \AND
       \Name{Alexander Vinogradov}
       \Email{vinogradov.am@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)
       \AND
       \Name{Daniil Merkulov}
       \Email{daniil.merkulov@phystech.edu}\\
       \addr Moscow Institute of Physics and Technology (MIPT)\\
       Skoltech, HSE, AI4Science
       \AND
       \Name{Ivan Oseledets}
       \Email{i.oseledets@skoltech.ru}\\
       \addr AIRI, Skoltech\\
       Marchuk Institute of Numerical Mathematics}

@@

Unlike Schatten norms, which satisfy $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$ (for $p^{-1} + q^{-1} = 1$) and are thus closed under dualization, Ky Fan norms are generally not. Only two exceptional cases exist: the dual of the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa.

Thus working with duals of Ky Fan norms can open new possibilities:

@@

\paragraph{Connection to existing algorithms}
The Fanion family unifies several known algorithms as special cases:
\begin{itemize}
	\item \textbf{Neon} is Fanion-1 (rank-one updates)
	\item \textbf{Muon} is Fanion-$\min\{m,n\}$ (full-rank updates)
	\item \textbf{Dion} (unsharded): The rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without error feedback and without scaling of the update is actually Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written in a notation more similar to ours)
\end{itemize}

>>>>

\author{\Name{Alexey Kravatskiy}
       \Email{kravtskii.aiu@phystech.edu}\\
       \addr MIPT
       \AND
       \Name{Ivan Kozyrev}
       \Email{kozyrev.in@phystech.edu}\\
       \addr MIPT, INM RAS
       \AND
       \Name{Nikolai Kozlov}
       \Email{kozlov.na@phystech.edu}\\
       \addr MIPT
       \AND
       \Name{Alexander Vinogradov}
       \Email{vinogradov.am@phystech.edu}\\
       \addr MIPT
       \AND
       \Name{Daniil Merkulov}
       \Email{daniil.merkulov@phystech.edu}\\
       \addr MIPT, Skoltech, HSE, AI4Science
       \AND
       \Name{Ivan Oseledets}
       \Email{i.oseledets@skoltech.ru}\\
       \addr AIRI, Skoltech, INM RAS
}

@@

Unlike Schatten norms, which satisfy $\norm{\mM^t}_{S_p}^\dagger = \norm{\mM^t}_{S_q}$ (for $p^{-1} + q^{-1} = 1$) and are thus closed under dualization, Ky Fan norms are generally not. Only two exceptional cases exist: the dual of the Ky Fan $1$-norm $\norm{\cdot}_{\text{KF-}1}$ (the spectral norm) is the Ky Fan $\min(m,n)$-norm $\norm{\cdot}_{\text{KF-}\min(m,n)}$ (the nuclear norm), and vice versa. Thus working with duals of Ky Fan norms can open up new possibilities:

@@

\paragraph{Connection to existing algorithms}
The Fanion family unifies several known algorithms:
\begin{itemize}
	\item \textbf{Neon} is Fanion-1 (rank-one updates)
	\item \textbf{Muon} is Fanion-$\min\{m,n\}$ (full-rank updates)
	\item \textbf{Dion} (unsharded): The rank-$r$ Dion (Algorithm 1 from \citet{ahn2025dion}) without error feedback and without scaling of the update is actually Fanion-$r$ (see \citet{pethick2025understandingdion}, where Dion is written in a notation more similar to ours)
\end{itemize}
